---
title: "Applying HW3 to house data"
author: "Jacob Bayer"
date: "4/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(glmnet)
library(dplyr)
rm(list = ls())    #delete objects
cat("\014")
setwd("C:/Users/Jacob-Windows/OneDrive - The City University of New York/School/Spring 21/STA 9890/house-prices-advanced-regression-techniques")
house<-read.csv("train.csv", stringsAsFactors = TRUE)
```


```{r}
house = house %>% 
  purrr::discard(~sum(is.na(.x))/length(.x)* 100 >= 0.001)

house$HouseAge<-as.integer(format(Sys.Date(), "%Y"))-house$YearBuilt
house$RemodelAge<-as.integer(format(Sys.Date(), "%Y"))-house$YearRemodAdd
house$YrSinceSold<-as.integer(format(Sys.Date(), "%Y"))-house$YrSold

house$SalePriceLog <- log(house$SalePrice/1000)

SalePrice <- house$SalePrice

dropcols<-c('YrSold','YearBuilt','YearRemodAdd','SalePrice')
house <- house %>% select(-one_of(dropcols))

```

Here p is the number of variables remaining in the house data after some columns were removed. n is the number of observations in the house data. X is a matrisx
```{r}
set.seed(0)
X = model.matrix(SalePriceLog∼., house)
y = as.numeric(as.matrix(house['SalePriceLog']))
p = dim(X)[2]
n = dim(X)[1]
K = 10
d = ceiling(n/K)
set.seed(0)
i.mix = sample(1:n)
```

What should lam.net be?

```{r}
# Tuning parameter values for lasso, and ridge regression
lam.las = c(seq(1e-3,0.1,length=100),seq(0.12,2.5,length=100)) 
lam.rid = lam.las*1000
lam.net = lam.las*500
```


```{r}
nlam = length(lam.las)
# These two matrices store the prediction errors for each
# observation (along the rows), when we fit the model using
# each value of the tuning parameter (along the columns)
e.rid = matrix(0,n,nlam)
e.las = matrix(0,n,nlam)
e.net = matrix(0,n,nlam)
```


# Part A

```{r}
for (k in 1:K) {
  # Here there are K = 10 folds. D is n/k.
  cat("Fold",k,"\n")
  
  # i.mix is a shuffled index of X
  # The folds range endpoints:
  
  folds_range_start = (1+(k-1)*d)     # for (k in 1:K) {print(1+(k-1)*d)}
  folds_range_end = (k*d)             # for (k in 1:K) {print(k*d)}

    
  folds = folds_range_start:folds_range_end; # for (k in 1:K) {print((1+(k-1)*d):(k*d))}
  
  
  # For the first fold, the first 130 observations become the test data. For the second fold, the second
  # 130 observations become the test data, and so on.
  i.tr=i.mix[-folds]
  i.val=i.mix[folds]

  
  X.tr = X[i.tr,] # houseing predictors
  y.tr = y[i.tr]   # houseing responses
  X.val = X[i.val,] # validation predictors
  y.val = y[i.val]  # validation responses
  
  
  # TODO
  # Now use the function glmnet on the houseing data to get the 
  # ridge regression solutions at all tuning parameter values in
  # lam.rid, and the lasso solutions at all tuning parameter 
  # values in lam.las
  a.rid = glmnet(alpha = 0,x =  X.tr, y = y.tr, lambda = lam.rid) # for the ridge regression solutions, use alpha=0
  a.las = glmnet(alpha = 1, x =  X.tr, y = y.tr, lambda = lam.las) # for the lasso solutions, use alpha=1
  a.net = glmnet(alpha = 0.5, x =  X.tr, y = y.tr, lambda = lam.net)
  
  # Here we're actually going to reverse the column order of the
  # a.rid$beta and a.las$beta matrices, because we want their columns
  # to correspond to increasing lambda values (glmnet's default makes
  # it so that these are actually in decreasing lambda order), i.e.,
  # in the same order as our lam.rid and lam.las vectors
  rid.beta = as.matrix(a.rid$beta[,nlam:1])
  las.beta = as.matrix(a.las$beta[,nlam:1])
  net.beta = as.matrix(a.las$beta[,nlam:1])
  
  yhat.rid = X.val%*%rid.beta
  yhat.las = X.val%*%las.beta
  yhat.net = X.val%*%net.beta
  
  # The length of e.rid nlam*length(bstar)
  e.rid[i.val,] = (yhat.rid-y.val)^2
  e.las[i.val,] = (yhat.las-y.val)^2
  e.net[i.val,] = (yhat.net-y.val)^2
}
```

```{r}
# TODO
# Here you need to compute: 
# -cv.rid, cv.las: vectors of length nlam, giving the cross-validation
#  errors for ridge regression and the lasso, across all values of the
#  tuning parameter
cv.rid = colMeans(e.rid)
cv.las = colMeans(e.las)
cv.net = colMeans(e.net)
```

```{r}

# -se.rid, se.las: vectors of length nlam, giving the standard errors
#  of the cross-validation estimates for ridge regression and the lasso, 
#  across all values of the tuning parameter

pe.rid = matrix(0,K,nlam)
pe.las = matrix(0,K,nlam)
pe.net = matrix(0,K,nlam)

# This calculates the error for each fold
for (k in 1:K) {
  cat("Fold",k,"\n")
  folds=(1+(k-1)*d):(k*d);
  i.val=i.mix[folds]
  
  # pe.rid and pe.las are both nlam long
  pe.rid[k,] = colMeans(e.rid[i.val,])
  pe.las[k,] = colMeans(e.las[i.val,])
  pe.net[k,] = colMeans(e.net[i.val,])
}


se.rid = apply(pe.rid,2,sd)/sqrt(K)
se.las = apply(pe.las,2,sd)/sqrt(K)
se.net = apply(pe.net,2,sd)/sqrt(K)
```

```{r}

# Usual rule for choosing lambda
i1.rid = which.min(cv.rid)
i1.las = which.min(cv.las)
i1.net = which.min(cv.net)

cv.rid.1se = cv.rid[i1.rid] + se.rid[i1.rid]
cv.las.1se = cv.las[i1.las]+se.las[i1.las]
cv.net.1se = cv.las[i1.net]+se.las[i1.net]

i2.rid = which(cv.rid < cv.rid.1se)
i2.rid = tail(i2.rid, n=1)


i2.las = which(cv.las < cv.las.1se)
i2.las = tail(i2.las, n=1)

i2.net = which(cv.net < cv.net.1se)
i2.net = tail(i2.net, n=1)

```



```{r}
min.cv.rid=cv.rid[i1.rid]
min.cv.las=cv.las[i1.las]
min.cv.net=cv.las[i1.net]
print('Ridge Error:')
min.cv.rid
print("Lasso Error:")
min.cv.las
print("Elastic Net Error:")
min.cv.net
```

The lambda values that minimize cross validation mean square error are printed below:

```{r}
# Ridge
lambda.rid = lam.rid[i1.rid]

# Lasso
lambda.las = lam.las[i1.las]

# Elastic Net
lambda.net = lam.net[i1.net]

print('Ridge Lambda:')
lambda.rid
print("Lasso Lambda:")
lambda.las
print("Elastic Net Lambda:")
lambda.net
```



```{r}
a.rid = glmnet(X, y, lambda = lambda.rid, alpha=0)
a.las = glmnet(X, y, lambda = lambda.las, alpha=1) 
a.net = glmnet(X, y, lambda = lambda.net, alpha=0.5) 

rid.beta = as.matrix(a.rid$beta[,1])
las.beta = as.matrix(a.las$beta[,1]) 
net.beta = as.matrix(a.las$beta[,1]) 

yhat.rid = X%*%rid.beta
yhat.las = X%*%las.beta
yhat.net = X%*%las.beta

```



# SSE

```{r}
# Ridge SSE
sum((y-yhat.rid)^2)

# Lasso SSE
sum((y-yhat.las)^2)

# Lasso SSE
sum((y-yhat.net)^2)
```
# MSE

This seems good. These are low. 

```{r}
# Ridge SSE
mean((y-yhat.rid)^2)

# Lasso SSE
mean((y-yhat.las)^2)

# Lasso SSE
mean((y-yhat.net)^2)
```

I don't know if these are low

```{r}
# Ridge absolute error
sum(abs((y-yhat.rid)))

# Lasso absolute error
sum(abs((y-yhat.las)))

# Lasso absolute error
sum(abs((y-yhat.net)))
```
You should be able to convert these logarithm predictions back into sale price but there is a lot of error introduced there.

```{r}
# Ridge absolute error
sum(abs((SalePrice-exp(yhat.rid)*1000)))

# Lasso absolute error
sum(abs((SalePrice-exp(yhat.las)*1000)))

# Lasso absolute error
sum(abs((SalePrice-exp(yhat.net)*1000)))
```

```{r}
# Ridge SSE
mean((SalePrice-exp(yhat.rid)*1000)^2)

# Lasso SSE
mean((SalePrice-exp(yhat.las)*1000)^2)

# Lasso SSE
mean((SalePrice-exp(yhat.net)*1000)^2)
```
